---
title: "vaccinations-UK-first-dose-demographics"
author: "Vanessa Fillis"
date: "10/06/2021"
output: html_document
---

# Cumulative figures of those who have had a first dose broken down by age: a) 60+, b) under 30 c) 30-59

## Activating packages

```{r}
library(jsonlite)
library(httr)
library(dplyr)
library(zoo)
library(tidyr)
library(purrr)
```

## Importing the data 

```{r}
#' Extracts paginated data by requesting all of the pages
#' and combining the results.
#'
#' @param filters    API filters. See the API documentations for 
#'                   additional information.
#'                   
#' @param structure  Structure parameter. See the API documentations 
#'                   for additional information.
#'                   
#' @return list      Comprehensive list of dictionaries containing all 
#'                   the data for the given ``filter`` and ``structure`.`
get_paginated_data <- function (filters, structure) {
  
    endpoint     <- "https://api.coronavirus.data.gov.uk/v1/data"
    results      <- list()
    current_page <- 1
    
    repeat {
        httr::GET(
            url   = endpoint,
            query = list(
                filters   = paste(filters, collapse = ";"),
                structure = jsonlite::toJSON(structure, auto_unbox = TRUE),
                page      = current_page
            ),
            timeout(10)
        ) -> response
        
        # Handle errors:
        if ( response$status_code >= 400 ) {
            err_msg = httr::http_status(response)
            stop(err_msg)
        } else if ( response$status_code == 204 ) {
            break
        }
        
        # Convert response from binary to JSON:
        json_text <- content(response, "text")
        dt        <- jsonlite::fromJSON(json_text)
        results   <- rbind(results, dt$data)
        
        if ( is.null( dt$pagination$`next` ) ){
            break
        }
    
        current_page <- current_page + 1;
    }
    
    return(results)
    
}
# Create filters:
query_filters <- c(
    "areaType=nation",
    "areaName=England"
)

# Create the structure as a list or a list of lists:
query_structure <- list(
    date       = "date", 
    name       = "areaName",
    #code       = "areaCode"
    vaccination.ages = "vaccinationsAgeDemographics"
    #cumPeopleVaccinatedFirstDoseByVaccinationDate
)

vaccinations.UK <- get_paginated_data(query_filters, query_structure)
list(
  "Shape"                = dim(vaccinations.UK),
  "Data (first 3 items)" = vaccinations.UK[0:3, 0:-1]
) -> report
print(report)

```

## Unnest the data frame

The data frame is nested, so we need to unnest it 

```{r}
vaccinations.UK <- vaccinations.UK %>%
  unnest(vaccination.ages)
```

## Delete metrics we don't need 

Only want to look at the cumulative figures of those who have had a first dose

```{r}
vaccinations.UK <- vaccinations.UK[-c(4:6,8:13)]
```


## Combine age groups 

We want to look at three age groups:

a) 60+
b) under 30 
c) 30-59

```{r}
vaccinations.UK.under30 <- subset(vaccinations.UK, vaccinations.UK$age == "18_24" | vaccinations.UK$age == "25_29")

vaccinations.UK.under30 <- vaccinations.UK.under30 %>%
  group_by(date) %>%
  summarise(under30 = sum(cumPeopleVaccinatedFirstDoseByVaccinationDate))

vaccinations.UK.30to59 <- subset(vaccinations.UK, vaccinations.UK$age == "30_34" | vaccinations.UK$age == "35_39" | vaccinations.UK$age == "40_44" | vaccinations.UK$age == "45_49" | vaccinations.UK$age == "50_54" | vaccinations.UK$age == "55_59")

vaccinations.UK.30to59 <- vaccinations.UK.30to59 %>%
  group_by(date) %>%
  summarise(v30to59 = sum(cumPeopleVaccinatedFirstDoseByVaccinationDate))

vaccinations.UK.over60 <- subset(vaccinations.UK, vaccinations.UK$age == "60_64" | vaccinations.UK$age == "65_69" | vaccinations.UK$age == "70_74" | vaccinations.UK$age == "75_79" | vaccinations.UK$age == "80_84" | vaccinations.UK$age == "85_89" | vaccinations.UK$age == "90+")

vaccinations.UK.over60 <- vaccinations.UK.over60 %>%
  group_by(date) %>%
  summarise(over60 = sum(cumPeopleVaccinatedFirstDoseByVaccinationDate))

```

## Merge the three frames

```{r}
vaccinations.UK.groups <- merge(merge(vaccinations.UK.under30, vaccinations.UK.30to59, by = "date"), vaccinations.UK.over60, by = "date")
```

## Export 

```{r}
write.csv(vaccinations.UK.groups, file = "vaccinations-UK-first-dose-demographics.csv")
```

