---
title: "UK Covid cases"
author: "Vanessa Fillis"
date: "02/06/2021"
output: html_document
---

# Daily Covid cases in the UK 

## Activate packages

```{r}
library(jsonlite)
library(httr)
library(dplyr)
library(zoo)
```

## Getting the data 

Cases by date reported

```{r}
#' Extracts paginated data by requesting all of the pages
#' and combining the results.
#'
#' @param filters    API filters. See the API documentations for 
#'                   additional information.
#'                   
#' @param structure  Structure parameter. See the API documentations 
#'                   for additional information.
#'                   
#' @return list      Comprehensive list of dictionaries containing all 
#'                   the data for the given ``filter`` and ``structure`.`
get_paginated_data <- function (filters, structure) {
  
    endpoint     <- "https://api.coronavirus.data.gov.uk/v1/data"
    results      <- list()
    current_page <- 1
    
    repeat {
        httr::GET(
            url   = endpoint,
            query = list(
                filters   = paste(filters, collapse = ";"),
                structure = jsonlite::toJSON(structure, auto_unbox = TRUE),
                page      = current_page
            ),
            timeout(10)
        ) -> response
        
        # Handle errors:
        if ( response$status_code >= 400 ) {
            err_msg = httr::http_status(response)
            stop(err_msg)
        } else if ( response$status_code == 204 ) {
            break
        }
        
        # Convert response from binary to JSON:
        json_text <- content(response, "text")
        dt        <- jsonlite::fromJSON(json_text)
        results   <- rbind(results, dt$data)
        
        if ( is.null( dt$pagination$`next` ) ){
            break
        }
    
        current_page <- current_page + 1;
    }
    
    return(results)
    
}
# Create filters:
query_filters <- c(
    "areaType=nation"
)

# Create the structure as a list or a list of lists:
query_structure <- list(
    date       = "date", 
    name       = "areaName", 
    code       = "areaCode", 
    daily     = "newCasesByPublishDate",
    cumulative = "cumCasesByPublishDate"
)
cases <- get_paginated_data(query_filters, query_structure)
list(
  "Shape"                = dim(cases),
  "Data (first 3 items)" = cases[0:3, 0:-1]
) -> report
print(report)
#Rename columns for better distinction
cases <- cases %>%
  rename(
    daily.cases.specimen = daily,
    cumulative.cases.specimen = cumulative
  )
```

## Summarise for UK

```{r}
cases.UK <- cases %>%
  group_by(name) %>%
  group_by(date) %>%
  summarise(sum.daily.cases = sum(daily.cases.specimen))
```

## Calculating 7 day average

```{r}
cases.UK <- cases.UK %>%
  dplyr::arrange(desc(date)) %>% 
  dplyr::mutate(cases_07da = zoo::rollmean(sum.daily.cases, k = 7, align="left", fill = NA)) 
```

## Export UK

```{r}
write.csv(cases.UK, file = "cases-UK.csv")
```

## Filter for England

```{r}
cases.England <- subset(cases, cases$name == "England")
```

## Calculating 7 day average

```{r}
cases.England <- cases.England %>%
  #dplyr::arrange(desc(date)) %>% 
  dplyr::mutate(cases_07da = zoo::rollmean(daily.cases.specimen, k = 7, align="left", fill = NA)) 
```

## Export England

```{r}
write.csv(cases.England, file = "cases-England.csv")
```

## Filter for Scotland

```{r}
cases.Scotland <- subset(cases, cases$name == "Scotland")
```

## Calculating 7 day average

```{r}
cases.Scotland <- cases.Scotland %>%
  #dplyr::arrange(desc(date)) %>% 
  dplyr::mutate(cases_07da = zoo::rollmean(daily.cases.specimen, k = 7, align="left", fill = NA)) 
```

## Export Scotland

```{r}
write.csv(cases.Scotland, file = "cases-Scotland.csv")
```

## Filter for Wales

```{r}
cases.Wales <- subset(cases, cases$name == "Wales")
```

## Calculating 7 day average

```{r}
cases.Wales <- cases.Wales %>%
  #dplyr::arrange(desc(date)) %>% 
  dplyr::mutate(cases_07da = zoo::rollmean(daily.cases.specimen, k = 7, align="left", fill = NA)) 
```

## Export Wales

```{r}
write.csv(cases.Wales, file = "cases-Wales.csv")
```

## Filter for Northern Ireland

```{r}
cases.NI <- subset(cases, cases$name == "Northern Ireland")
```

## Calculating 7 day average

```{r}
cases.NI <- cases.NI %>%
  #dplyr::arrange(desc(date)) %>% 
  dplyr::mutate(cases_07da = zoo::rollmean(daily.cases.specimen, k = 7, align="left", fill = NA)) 
```

## Export Northern Ireland

```{r}
write.csv(cases.NI, file = "cases-Northern-Ireland.csv")
```

