---
title: "API test"
author: "Vanessa Fillis"
date: "10/11/2020"
output: html_document
---

# Download COVID-data with an API

Downloading the data from https://coronavirus.data.gov.uk using an API. 

At first, activating necessary packages. 

```{r}
library(jsonlite)
library(httr)
library(dplyr)
```

## Cases England

Cases by specimen date

```{r}
#' Extracts paginated data by requesting all of the pages
#' and combining the results.
#'
#' @param filters    API filters. See the API documentations for 
#'                   additional information.
#'                   
#' @param structure  Structure parameter. See the API documentations 
#'                   for additional information.
#'                   
#' @return list      Comprehensive list of dictionaries containing all 
#'                   the data for the given ``filter`` and ``structure`.`
get_paginated_data <- function (filters, structure) {
  
    endpoint     <- "https://api.coronavirus.data.gov.uk/v1/data"
    results      <- list()
    current_page <- 1
    
    repeat {

        httr::GET(
            url   = endpoint,
            query = list(
                filters   = paste(filters, collapse = ";"),
                structure = jsonlite::toJSON(structure, auto_unbox = TRUE),
                page      = current_page
            ),
            timeout(10)
        ) -> response
        
        # Handle errors:
        if ( response$status_code >= 400 ) {
            err_msg = httr::http_status(response)
            stop(err_msg)
        } else if ( response$status_code == 204 ) {
            break
        }
        
        # Convert response from binary to JSON:
        json_text <- content(response, "text")
        dt        <- jsonlite::fromJSON(json_text)
        results   <- rbind(results, dt$data)
        
        if ( is.null( dt$pagination$`next` ) ){
            break
        }
    
        current_page <- current_page + 1;
    }
    
    return(results)
    
}

# Create filters:
query_filters <- c(
    "areaType=nation",
    "areaName=England"
)

# Create the structure as a list or a list of lists:
query_structure <- list(
    date       = "date", 
    name       = "areaName", 
    code       = "areaCode", 
    daily      = "newCasesBySpecimenDate",
    cumulative = "cumCasesBySpecimenDate"
)

casesEnglandSpecimen <- get_paginated_data(query_filters, query_structure)

list(
  "Shape"                = dim(casesEnglandSpecimen),
  "Data (first 3 items)" = casesEnglandSpecimen[0:3, 0:-1]
) -> report

print(report)

#Rename columns for better distinction
casesEnglandSpecimen <- casesEnglandSpecimen %>%
  rename(
    daily.cases.specimen = daily,
    cumulative.cases.specimen = cumulative
  )
```

## Cases England

Cases by date reported

```{r}
#' Extracts paginated data by requesting all of the pages
#' and combining the results.
#'
#' @param filters    API filters. See the API documentations for 
#'                   additional information.
#'                   
#' @param structure  Structure parameter. See the API documentations 
#'                   for additional information.
#'                   
#' @return list      Comprehensive list of dictionaries containing all 
#'                   the data for the given ``filter`` and ``structure`.`
get_paginated_data <- function (filters, structure) {
  
    endpoint     <- "https://api.coronavirus.data.gov.uk/v1/data"
    results      <- list()
    current_page <- 1
    
    repeat {

        httr::GET(
            url   = endpoint,
            query = list(
                filters   = paste(filters, collapse = ";"),
                structure = jsonlite::toJSON(structure, auto_unbox = TRUE),
                page      = current_page
            ),
            timeout(10)
        ) -> response
        
        # Handle errors:
        if ( response$status_code >= 400 ) {
            err_msg = httr::http_status(response)
            stop(err_msg)
        } else if ( response$status_code == 204 ) {
            break
        }
        
        # Convert response from binary to JSON:
        json_text <- content(response, "text")
        dt        <- jsonlite::fromJSON(json_text)
        results   <- rbind(results, dt$data)
        
        if ( is.null( dt$pagination$`next` ) ){
            break
        }
    
        current_page <- current_page + 1;
    }
    
    return(results)
    
}

# Create filters:
query_filters <- c(
    "areaType=nation",
    "areaName=England"
)

# Create the structure as a list or a list of lists:
query_structure <- list(
    date       = "date", 
    name       = "areaName", 
    code       = "areaCode", 
    daily     = "newCasesByPublishDate",
    cumulative = "cumCasesByPublishDate"
)

casesEnglandReported <- get_paginated_data(query_filters, query_structure)

list(
  "Shape"                = dim(casesEnglandReported),
  "Data (first 3 items)" = casesEnglandReported[0:3, 0:-1]
) -> report

print(report)

#Rename columns for better distinction
casesEnglandReported <- casesEnglandReported %>%
  rename(
    daily.cases.reported = daily,
    cumulative.cases.reported = cumulative
  )
```

## Cases Regions 

Cases by specimen date (regional breakdown of cases is only available by specimen date)

```{r}
#' Extracts paginated data by requesting all of the pages
#' and combining the results.
#'
#' @param filters    API filters. See the API documentations for 
#'                   additional information.
#'                   
#' @param structure  Structure parameter. See the API documentations 
#'                   for additional information.
#'                   
#' @return list      Comprehensive list of dictionaries containing all 
#'                   the data for the given ``filter`` and ``structure`.`
get_paginated_data <- function (filters, structure) {
  
    endpoint     <- "https://api.coronavirus.data.gov.uk/v1/data"
    results      <- list()
    current_page <- 1
    
    repeat {

        httr::GET(
            url   = endpoint,
            query = list(
                filters   = paste(filters, collapse = ";"),
                structure = jsonlite::toJSON(structure, auto_unbox = TRUE),
                page      = current_page
            ),
            timeout(10)
        ) -> response
        
        # Handle errors:
        if ( response$status_code >= 400 ) {
            err_msg = httr::http_status(response)
            stop(err_msg)
        } else if ( response$status_code == 204 ) {
            break
        }
        
        # Convert response from binary to JSON:
        json_text <- content(response, "text")
        dt        <- jsonlite::fromJSON(json_text)
        results   <- rbind(results, dt$data)
        
        if ( is.null( dt$pagination$`next` ) ){
            break
        }
        
        current_page <- current_page + 1;

    }
    
    return(results)
    
}

# Create filters:
query_filters <- c(
    "areaType=region"
)

# Create the structure as a list or a list of lists:
query_structure <- list(
    date       = "date", 
    name       = "areaName", 
    code       = "areaCode", 
    daily      = "newCasesBySpecimenDate",
    cumulative = "cumCasesBySpecimenDate"
)

casesRegions <- get_paginated_data(query_filters, query_structure)

list(
  "Shape"                = dim(casesRegions),
  "Data (first 3 items)" = casesRegions[0:3, 0:-1]
) -> report

print(report)

#Rename columns for better distinction
casesRegions <- casesRegions %>%
  rename(
    daily.cases = daily,
    cumulative.cases = cumulative
  )
```

## Cases Upper-tier local authorities

Cases by specimen date

```{r}
#' Extracts paginated data by requesting all of the pages
#' and combining the results.
#'
#' @param filters    API filters. See the API documentations for 
#'                   additional information.
#'                   
#' @param structure  Structure parameter. See the API documentations 
#'                   for additional information.
#'                   
#' @return list      Comprehensive list of dictionaries containing all 
#'                   the data for the given ``filter`` and ``structure`.`
get_paginated_data <- function (filters, structure) {
  
    endpoint     <- "https://api.coronavirus.data.gov.uk/v1/data"
    results      <- list()
    current_page <- 1
    
    repeat {

        httr::GET(
            url   = endpoint,
            query = list(
                filters   = paste(filters, collapse = ";"),
                structure = jsonlite::toJSON(structure, auto_unbox = TRUE),
                page      = current_page
            ),
            timeout(10)
        ) -> response
        
        # Handle errors:
        if ( response$status_code >= 400 ) {
            err_msg = httr::http_status(response)
            stop(err_msg)
        } else if ( response$status_code == 204 ) {
            break
        }
        
        # Convert response from binary to JSON:
        json_text <- content(response, "text")
        dt        <- jsonlite::fromJSON(json_text)
        results   <- rbind(results, dt$data)
        
        if ( is.null( dt$pagination$`next` ) ){
            break
        }
        
        current_page <- current_page + 1;

    }
    
    return(results)
    
}

# Create filters:
query_filters <- c(
    "areaType=utla"
)

# Create the structure as a list or a list of lists:
query_structure <- list(
    date       = "date", 
    name       = "areaName", 
    code       = "areaCode", 
    daily      = "newCasesBySpecimenDate",
    cumulative = "cumCasesBySpecimenDate"
)

casesUpperLA <- get_paginated_data(query_filters, query_structure)

list(
  "Shape"                = dim(casesUpperLA),
  "Data (first 3 items)" = casesUpperLA[0:3, 0:-1]
) -> report

print(report)

# Rename columns for better distinction
casesUpperLA <- casesUpperLA %>%
  rename(
    daily.cases = daily,
    cumulative.cases = cumulative
  )
```

## Cases Lower-tier local authorities

Cases by specimen date

```{r}
#' Extracts paginated data by requesting all of the pages
#' and combining the results.
#'
#' @param filters    API filters. See the API documentations for 
#'                   additional information.
#'                   
#' @param structure  Structure parameter. See the API documentations 
#'                   for additional information.
#'                   
#' @return list      Comprehensive list of dictionaries containing all 
#'                   the data for the given ``filter`` and ``structure`.`
get_paginated_data <- function (filters, structure) {
  
    endpoint     <- "https://api.coronavirus.data.gov.uk/v1/data"
    results      <- list()
    current_page <- 1
    
    repeat {

        httr::GET(
            url   = endpoint,
            query = list(
                filters   = paste(filters, collapse = ";"),
                structure = jsonlite::toJSON(structure, auto_unbox = TRUE),
                page      = current_page
            ),
            timeout(10)
        ) -> response
        
        # Handle errors:
        if ( response$status_code >= 400 ) {
            err_msg = httr::http_status(response)
            stop(err_msg)
        } else if ( response$status_code == 204 ) {
            break
        }
        
        # Convert response from binary to JSON:
        json_text <- content(response, "text")
        dt        <- jsonlite::fromJSON(json_text)
        results   <- rbind(results, dt$data)
        
        if ( is.null( dt$pagination$`next` ) ){
            break
        }
        
        current_page <- current_page + 1;

    }
    
    return(results)
    
}


# Create filters:
query_filters <- c(
    "areaType=ltla"
)

# Create the structure as a list or a list of lists:
query_structure <- list(
    date       = "date", 
    name       = "areaName", 
    code       = "areaCode", 
    daily      = "newCasesBySpecimenDate",
    cumulative = "cumCasesBySpecimenDate"
)

casesLowerLA <- get_paginated_data(query_filters, query_structure)

list(
  "Shape"                = dim(casesLowerLA),
  "Data (first 3 items)" = casesLowerLA[0:3, 0:-1]
) -> report

print(report)

# Rename columns for better distinction
casesLowerLA <- casesLowerLA %>%
  rename(
    daily.cases = daily,
    cumulative.cases = cumulative
  )
```

## Deaths England

Deaths within 28 days of positive test by death date

```{r}
#' Extracts paginated data by requesting all of the pages
#' and combining the results.
#'
#' @param filters    API filters. See the API documentations for 
#'                   additional information.
#'                   
#' @param structure  Structure parameter. See the API documentations 
#'                   for additional information.
#'                   
#' @return list      Comprehensive list of dictionaries containing all 
#'                   the data for the given ``filter`` and ``structure`.`
get_paginated_data <- function (filters, structure) {
  
    endpoint     <- "https://api.coronavirus.data.gov.uk/v1/data"
    results      <- list()
    current_page <- 1
    
    repeat {

        httr::GET(
            url   = endpoint,
            query = list(
                filters   = paste(filters, collapse = ";"),
                structure = jsonlite::toJSON(structure, auto_unbox = TRUE),
                page      = current_page
            ),
            timeout(10)
        ) -> response
        
        # Handle errors:
        if ( response$status_code >= 400 ) {
            err_msg = httr::http_status(response)
            stop(err_msg)
        } else if ( response$status_code == 204 ) {
            break
        }
        
        # Convert response from binary to JSON:
        json_text <- content(response, "text")
        dt        <- jsonlite::fromJSON(json_text)
        results   <- rbind(results, dt$data)
        
        if ( is.null( dt$pagination$`next` ) ){
            break
        }
    
        current_page <- current_page + 1;
    }
    
    return(results)
    
}

# Create filters:
query_filters <- c(
    "areaType=nation",
    "areaName=England"
)

# Create the structure as a list or a list of lists:
query_structure <- list(
    date       = "date", 
    name       = "areaName", 
    code       = "areaCode", 
    daily      = "newDeaths28DaysByDeathDate",
    cumulative = "cumDeaths28DaysByDeathDate"
)

deathsEngland <- get_paginated_data(query_filters, query_structure)

list(
  "Shape"                = dim(deathsEngland),
  "Data (first 3 items)" = deathsEngland[0:3, 0:-1]
) -> report

print(report)

#Rename columns for better distinction
deathsEngland <- deathsEngland %>%
  rename(
    daily.deaths = daily,
    cumulative.deaths = cumulative
  )
```

## Deaths Regions

Deaths within 28 days of positive test by death date

```{r}
#' Extracts paginated data by requesting all of the pages
#' and combining the results.
#'
#' @param filters    API filters. See the API documentations for 
#'                   additional information.
#'                   
#' @param structure  Structure parameter. See the API documentations 
#'                   for additional information.
#'                   
#' @return list      Comprehensive list of dictionaries containing all 
#'                   the data for the given ``filter`` and ``structure`.`
get_paginated_data <- function (filters, structure) {
  
    endpoint     <- "https://api.coronavirus.data.gov.uk/v1/data"
    results      <- list()
    current_page <- 1
    
    repeat {

        httr::GET(
            url   = endpoint,
            query = list(
                filters   = paste(filters, collapse = ";"),
                structure = jsonlite::toJSON(structure, auto_unbox = TRUE),
                page      = current_page
            ),
            timeout(10)
        ) -> response
        
        # Handle errors:
        if ( response$status_code >= 400 ) {
            err_msg = httr::http_status(response)
            stop(err_msg)
        } else if ( response$status_code == 204 ) {
            break
        }
        
        # Convert response from binary to JSON:
        json_text <- content(response, "text")
        dt        <- jsonlite::fromJSON(json_text)
        results   <- rbind(results, dt$data)
        
        if ( is.null( dt$pagination$`next` ) ){
            break
        }
    
        current_page <- current_page + 1;
    }
    
    return(results)
    
}

# Create filters:
query_filters <- c(
    "areaType=region"
)

# Create the structure as a list or a list of lists:
query_structure <- list(
    date       = "date", 
    name       = "areaName", 
    code       = "areaCode", 
    daily      = "newDeaths28DaysByDeathDate",
    cumulative = "cumDeaths28DaysByDeathDate"
)

deathsRegions <- get_paginated_data(query_filters, query_structure)

list(
  "Shape"                = dim(deathsRegions),
  "Data (first 3 items)" = deathsRegions[0:3, 0:-1]
) -> report

print(report)

#Rename columns for better distinction
deathsRegions <- deathsRegions %>%
  rename(
    daily.deaths = daily,
    cumulative.deaths = cumulative
  )
```

## Deaths Upper-tier local authorities

Deaths within 28 days of positive test by death date

```{r}
#' Extracts paginated data by requesting all of the pages
#' and combining the results.
#'
#' @param filters    API filters. See the API documentations for 
#'                   additional information.
#'                   
#' @param structure  Structure parameter. See the API documentations 
#'                   for additional information.
#'                   
#' @return list      Comprehensive list of dictionaries containing all 
#'                   the data for the given ``filter`` and ``structure`.`
get_paginated_data <- function (filters, structure) {
  
    endpoint     <- "https://api.coronavirus.data.gov.uk/v1/data"
    results      <- list()
    current_page <- 1
    
    repeat {

        httr::GET(
            url   = endpoint,
            query = list(
                filters   = paste(filters, collapse = ";"),
                structure = jsonlite::toJSON(structure, auto_unbox = TRUE),
                page      = current_page
            ),
            timeout(10)
        ) -> response
        
        # Handle errors:
        if ( response$status_code >= 400 ) {
            err_msg = httr::http_status(response)
            stop(err_msg)
        } else if ( response$status_code == 204 ) {
            break
        }
        
        # Convert response from binary to JSON:
        json_text <- content(response, "text")
        dt        <- jsonlite::fromJSON(json_text)
        results   <- rbind(results, dt$data)
        
        if ( is.null( dt$pagination$`next` ) ){
            break
        }
    
        current_page <- current_page + 1;
    }
    
    return(results)
    
}

# Create filters:
query_filters <- c(
    "areaType=utla"
)

# Create the structure as a list or a list of lists:
query_structure <- list(
    date       = "date", 
    name       = "areaName", 
    code       = "areaCode", 
    daily     = "newDeaths28DaysByDeathDate",
    cumulative = "cumDeaths28DaysByDeathDate"
)

deathsUpperLA <- get_paginated_data(query_filters, query_structure)

list(
  "Shape"                = dim(deathsUpperLA),
  "Data (first 3 items)" = deathsUpperLA[0:3, 0:-1]
) -> report

print(report)

#Rename columns for better distinction
deathsUpperLA <- deathsUpperLA %>%
  rename(
    daily.deaths = daily,
    cumulative.deaths = cumulative
  )
```

## Deaths Lower-tier local authorities

Deaths within 28 days of positive test by death date

```{r}
#' Extracts paginated data by requesting all of the pages
#' and combining the results.
#'
#' @param filters    API filters. See the API documentations for 
#'                   additional information.
#'                   
#' @param structure  Structure parameter. See the API documentations 
#'                   for additional information.
#'                   
#' @return list      Comprehensive list of dictionaries containing all 
#'                   the data for the given ``filter`` and ``structure`.`
get_paginated_data <- function (filters, structure) {
  
    endpoint     <- "https://api.coronavirus.data.gov.uk/v1/data"
    results      <- list()
    current_page <- 1
    
    repeat {

        httr::GET(
            url   = endpoint,
            query = list(
                filters   = paste(filters, collapse = ";"),
                structure = jsonlite::toJSON(structure, auto_unbox = TRUE),
                page      = current_page
            ),
            timeout(10)
        ) -> response
        
        # Handle errors:
        if ( response$status_code >= 400 ) {
            err_msg = httr::http_status(response)
            stop(err_msg)
        } else if ( response$status_code == 204 ) {
            break
        }
        
        # Convert response from binary to JSON:
        json_text <- content(response, "text")
        dt        <- jsonlite::fromJSON(json_text)
        results   <- rbind(results, dt$data)
        
        if ( is.null( dt$pagination$`next` ) ){
            break
        }
    
        current_page <- current_page + 1;
    }
    
    return(results)
    
}

# Create filters:
query_filters <- c(
    "areaType=ltla"
)

# Create the structure as a list or a list of lists:
query_structure <- list(
    date       = "date", 
    name       = "areaName", 
    code       = "areaCode", 
    daily     = "newDeaths28DaysByDeathDate",
    cumulative = "cumDeaths28DaysByDeathDate"
)

deathsLowerLA <- get_paginated_data(query_filters, query_structure)

list(
  "Shape"                = dim(deathsLowerLA),
  "Data (first 3 items)" = deathsLowerLA[0:3, 0:-1]
) -> report

print(report)

#Rename columns for better distinction
deathsLowerLA <- deathsLowerLA %>%
  rename(
    daily.deaths = daily,
    cumulative.deaths = cumulative
  )
```

## Healthcare Data England

includes: 
- New admissions: Daily numbers of COVID-19 patients admitted to hospital
- COVID-19 occupied beds with mechanical ventilators: Confirmed COVID-19 patients in mechanical ventilation beds
- Hospital cases: Daily count of confirmed COVID-19 patients in hospital at midnight the preceding night

```{r}
#' Extracts paginated data by requesting all of the pages
#' and combining the results.
#'
#' @param filters    API filters. See the API documentations for 
#'                   additional information.
#'                   
#' @param structure  Structure parameter. See the API documentations 
#'                   for additional information.
#'                   
#' @return list      Comprehensive list of dictionaries containing all 
#'                   the data for the given ``filter`` and ``structure`.`
get_paginated_data <- function (filters, structure) {
  
    endpoint     <- "https://api.coronavirus.data.gov.uk/v1/data"
    results      <- list()
    current_page <- 1
    
    repeat {

        httr::GET(
            url   = endpoint,
            query = list(
                filters   = paste(filters, collapse = ";"),
                structure = jsonlite::toJSON(structure, auto_unbox = TRUE),
                page      = current_page
            ),
            timeout(10)
        ) -> response
        
        # Handle errors:
        if ( response$status_code >= 400 ) {
            err_msg = httr::http_status(response)
            stop(err_msg)
        } else if ( response$status_code == 204 ) {
            break
        }
        
        # Convert response from binary to JSON:
        json_text <- content(response, "text")
        dt        <- jsonlite::fromJSON(json_text)
        results   <- rbind(results, dt$data)
        
        if ( is.null( dt$pagination$`next` ) ){
            break
        }
    
        current_page <- current_page + 1;
    }
    
    return(results)
    
}

# Create filters:
query_filters <- c(
    "areaType=nation",
    "areaName=England"
)

# Create the structure as a list or a list of lists:
query_structure <- list(
    date       = "date", 
    name       = "areaName", 
    code       = "areaCode", 
    admissions = "newAdmissions",
    mvbeds     = "covidOccupiedMVBeds",
    hospcases  = "hospitalCases"
)

healthcareEngland <- get_paginated_data(query_filters, query_structure)

list(
  "Shape"                = dim(healthcareEngland),
  "Data (first 3 items)" = healthcareEngland[0:3, 0:-1]
) -> report

print(report)

#Rename columns for better distinction
healthcareEngland <- healthcareEngland %>%
  rename(
    daily.newAdmissions = admissions,
    daily.covidOccupiedMVBeds = mvbeds,
    daily.hospitalCases = hospcases
  )
```

## Healthcare Data NHS regions

includes:
- New admissions: Daily numbers of COVID-19 patients admitted to hospital
- COVID-19 occupied beds with mechanical ventilators: Confirmed COVID-19 patients in mechanical ventilation beds
- Hospital cases: Daily count of confirmed COVID-19 patients in hospital at midnight the preceding night

```{r}
#' Extracts paginated data by requesting all of the pages
#' and combining the results.
#'
#' @param filters    API filters. See the API documentations for 
#'                   additional information.
#'                   
#' @param structure  Structure parameter. See the API documentations 
#'                   for additional information.
#'                   
#' @return list      Comprehensive list of dictionaries containing all 
#'                   the data for the given ``filter`` and ``structure`.`
get_paginated_data <- function (filters, structure) {
  
    endpoint     <- "https://api.coronavirus.data.gov.uk/v1/data"
    results      <- list()
    current_page <- 1
    
    repeat {

        httr::GET(
            url   = endpoint,
            query = list(
                filters   = paste(filters, collapse = ";"),
                structure = jsonlite::toJSON(structure, auto_unbox = TRUE),
                page      = current_page
            ),
            timeout(10)
        ) -> response
        
        # Handle errors:
        if ( response$status_code >= 400 ) {
            err_msg = httr::http_status(response)
            stop(err_msg)
        } else if ( response$status_code == 204 ) {
            break
        }
        
        # Convert response from binary to JSON:
        json_text <- content(response, "text")
        dt        <- jsonlite::fromJSON(json_text)
        results   <- rbind(results, dt$data)
        
        if ( is.null( dt$pagination$`next` ) ){
            break
        }
    
        current_page <- current_page + 1;
    }
    
    return(results)
    
}

# Create filters:
query_filters <- c(
    "areaType=nhsRegion"
)

# Create the structure as a list or a list of lists:
query_structure <- list(
    date       = "date", 
    name       = "areaName", 
    code       = "areaCode", 
    admissions = "newAdmissions",
    mvbeds     = "covidOccupiedMVBeds",
    hospcases  = "hospitalCases"
)

healthcareNhsRegions <- get_paginated_data(query_filters, query_structure)

list(
  "Shape"                = dim(healthcareNhsRegions),
  "Data (first 3 items)" = healthcareNhsRegions[0:3, 0:-1]
) -> report

print(report)

#Rename columns for better distinction
healthcareNhsRegions <- healthcareNhsRegions %>%
  rename(
    daily.newAdmissions = admissions,
    daily.covidOccupiedMVBeds = mvbeds,
    daily.hospitalCases = hospcases
  )
```

## Testing England

Number of lab-confirmed positive, negative or void COVID-19 test results by testing pillar (see below).

The government's mass testing programme includes four types of tests known as pillars:

- Pillar 1 NHS and PHE Testing: Virus testing in Public Health England (PHE) labs and NHS hospitals for those with a clinical need, and health and care workers

- Pillar 2 UK Government testing programme: Virus testing for the wider population, as set out in government guidance 

- Pillar 3 Antibody testing: Antibody serology testing to show if people have antibodies from having had COVID-19, reported from 1st June onwards

- Pillar 4 Surveillance testing: Virus testing and antibody serology testing for national surveillance supported by PHE, ONS, Biobank, universities and other partners to learn more about the prevalence and spread of the virus and for other testing research purposes, for example on the accuracy and ease of use of home testing.

Note: Data are only available from 1 April 2020. Up to 13 July 2020, only pillar 1 tests were included. Pillar 2 tests are available from 14 July 2020.
Testing data is only available for nations (and not for regions or LAs). 
Data for surveillance testing (pillar 4) are only available for the UK as a whole.

```{r}
#' Extracts paginated data by requesting all of the pages
#' and combining the results.
#'
#' @param filters    API filters. See the API documentations for 
#'                   additional information.
#'                   
#' @param structure  Structure parameter. See the API documentations 
#'                   for additional information.
#'                   
#' @return list      Comprehensive list of dictionaries containing all 
#'                   the data for the given ``filter`` and ``structure`.`
get_paginated_data <- function (filters, structure) {
  
    endpoint     <- "https://api.coronavirus.data.gov.uk/v1/data"
    results      <- list()
    current_page <- 1
    
    repeat {

        httr::GET(
            url   = endpoint,
            query = list(
                filters   = paste(filters, collapse = ";"),
                structure = jsonlite::toJSON(structure, auto_unbox = TRUE),
                page      = current_page
            ),
            timeout(10)
        ) -> response
        
        # Handle errors:
        if ( response$status_code >= 400 ) {
            err_msg = httr::http_status(response)
            stop(err_msg)
        } else if ( response$status_code == 204 ) {
            break
        }
        
        # Convert response from binary to JSON:
        json_text <- content(response, "text")
        dt        <- jsonlite::fromJSON(json_text)
        results   <- rbind(results, dt$data)
        
        if ( is.null( dt$pagination$`next` ) ){
            break
        }
    
        current_page <- current_page + 1;
    }
    
    return(results)
    
}

# Create filters:
query_filters <- c(
    "areaType=nation",
    "areaName=England"
)

# Create the structure as a list or a list of lists:
query_structure <- list(
    date       = "date", 
    name       = "areaName", 
    code       = "areaCode", 
    pillarone  = "newPillarOneTestsByPublishDate",
    pillartwo  = "newPillarTwoTestsByPublishDate",
    pillarthree= "newPillarThreeTestsByPublishDate",
    pillarfour = "newPillarFourTestsByPublishDate",
    newtests   = "newTestsByPublishDate"
)

testingEngland <- get_paginated_data(query_filters, query_structure)

list(
  "Shape"                = dim(testingEngland),
  "Data (first 3 items)" = testingEngland[0:3, 0:-1]
) -> report

print(report)

#Rename columns for better distinction
testingEngland <- testingEngland %>%
  rename(
    daily.newTests.pillarOne = pillarone,
    daily.newTests.pillarTwo = pillartwo,
    daily.newTests.pillarThree = pillarthree,
    daily.newTests.pillarFour = pillarfour,
    daily.newTests = newtests  
  )
```


