---
title: "Total and positive tests in England"
author: "Vanessa Fillis"
date: "05/01/2021"
output: html_document
---

# Positive testing rate before the third lockdown

At first, we have to first import data on cases and testing capacity. 

Daily tests and cases can be found on the [website of the UK government]("https://coronavirus.data.gov.uk/details/deaths"). We will look at all tests in all pillars as well as cases by date reported.

Since the third lockdown isn't imposed UK-wide, but England, Scotland, Wales and Northern Ireland each have different restrictions, we will only look at data for England.

## Activate packages

At first, we will activate necessary packages. 

```{r}
library(jsonlite)
library(httr)
library(dplyr)
library(lubridate)
```

## Import the data

We will first import the data of all tests in England. 

## Testing England

Number of lab-confirmed positive, negative or void COVID-19 test results by testing pillar (see below).

The government's mass testing programme includes four types of tests known as pillars:

- Pillar 1 NHS and PHE Testing: Virus testing in Public Health England (PHE) labs and NHS hospitals for those with a clinical need, and health and care workers

- Pillar 2 UK Government testing programme: Virus testing for the wider population, as set out in government guidance 

- Pillar 3 Antibody testing: Antibody serology testing to show if people have antibodies from having had COVID-19, reported from 1st June onwards

- Pillar 4 Surveillance testing: Virus testing and antibody serology testing for national surveillance supported by PHE, ONS, Biobank, universities and other partners to learn more about the prevalence and spread of the virus and for other testing research purposes, for example on the accuracy and ease of use of home testing.

Note: Data is only available from 1 April 2020. Up to 13 July 2020, only pillar 1 tests were included. Pillar 2 tests are available from 14 July 2020.
Testing data is only available for nations (and not for regions or LAs). 
Data for surveillance testing (pillar 4) are only available for the UK as a whole.

```{r}
#' Extracts paginated data by requesting all of the pages
#' and combining the results.
#'
#' @param filters    API filters. See the API documentations for 
#'                   additional information.
#'                   
#' @param structure  Structure parameter. See the API documentations 
#'                   for additional information.
#'                   
#' @return list      Comprehensive list of dictionaries containing all 
#'                   the data for the given ``filter`` and ``structure`.`
get_paginated_data <- function (filters, structure) {
  
    endpoint     <- "https://api.coronavirus.data.gov.uk/v1/data"
    results      <- list()
    current_page <- 1
    
    repeat {

        httr::GET(
            url   = endpoint,
            query = list(
                filters   = paste(filters, collapse = ";"),
                structure = jsonlite::toJSON(structure, auto_unbox = TRUE),
                page      = current_page
            ),
            timeout(10)
        ) -> response
        
        # Handle errors:
        if ( response$status_code >= 400 ) {
            err_msg = httr::http_status(response)
            stop(err_msg)
        } else if ( response$status_code == 204 ) {
            break
        }
        
        # Convert response from binary to JSON:
        json_text <- content(response, "text")
        dt        <- jsonlite::fromJSON(json_text)
        results   <- rbind(results, dt$data)
        
        if ( is.null( dt$pagination$`next` ) ){
            break
        }
    
        current_page <- current_page + 1;
    }
    
    return(results)
    
}

# Create filters:
query_filters <- c(
    "areaType=nation",
    "areaName=England"
)

# Create the structure as a list or a list of lists:
query_structure <- list(
    date       = "date", 
    name       = "areaName", 
    code       = "areaCode", 
    pillarone  = "newPillarOneTestsByPublishDate",
    pillartwo  = "newPillarTwoTestsByPublishDate",
    pillarthree= "newPillarThreeTestsByPublishDate",
    pillarfour = "newPillarFourTestsByPublishDate",
    newtests   = "newTestsByPublishDate"
)

tests <- get_paginated_data(query_filters, query_structure)

list(
  "Shape"                = dim(tests),
  "Data (first 3 items)" = tests[0:3, 0:-1]
) -> report

print(report)

#Rename columns for better distinction
tests <- tests %>%
  rename(
    daily.newTests.pillarOne = pillarone,
    daily.newTests.pillarTwo = pillartwo,
    daily.newTests.pillarThree = pillarthree,
    daily.newTests.pillarFour = pillarfour,
    daily.newTests = newtests  
  )
```

## Calculate total tests for each day

The column "daily.newTests" shows the total tests on each day - but only from 13 July. Before the column is empty. 

We will therefore create a new column with the daily total tests since 31 March.

Before we can do that we will replace all NAs with 0. 

```{r}
tests$daily.newTests.pillarTwo <- tests$daily.newTests.pillarTwo %>% tidyr::replace_na(0)
tests$daily.newTests.pillarThree <- tests$daily.newTests.pillarThree %>% tidyr::replace_na(0)
tests$daily.newTests.pillarFour <- tests$daily.newTests.pillarFour %>% tidyr::replace_na(0)
```

Now we can calculate the totals. 

```{r}
tests$total.tests <- tests$daily.newTests.pillarOne+tests$daily.newTests.pillarTwo+tests$daily.newTests.pillarThree+tests$daily.newTests.pillarFour
```

## Import daily positive cases

To calculate the positive testing rate we need data on daily new cases. 

Cases by date reported

```{r}
#' Extracts paginated data by requesting all of the pages
#' and combining the results.
#'
#' @param filters    API filters. See the API documentations for 
#'                   additional information.
#'                   
#' @param structure  Structure parameter. See the API documentations 
#'                   for additional information.
#'                   
#' @return list      Comprehensive list of dictionaries containing all 
#'                   the data for the given ``filter`` and ``structure`.`
get_paginated_data <- function (filters, structure) {
  
    endpoint     <- "https://api.coronavirus.data.gov.uk/v1/data"
    results      <- list()
    current_page <- 1
    
    repeat {

        httr::GET(
            url   = endpoint,
            query = list(
                filters   = paste(filters, collapse = ";"),
                structure = jsonlite::toJSON(structure, auto_unbox = TRUE),
                page      = current_page
            ),
            timeout(10)
        ) -> response
        
        # Handle errors:
        if ( response$status_code >= 400 ) {
            err_msg = httr::http_status(response)
            stop(err_msg)
        } else if ( response$status_code == 204 ) {
            break
        }
        
        # Convert response from binary to JSON:
        json_text <- content(response, "text")
        dt        <- jsonlite::fromJSON(json_text)
        results   <- rbind(results, dt$data)
        
        if ( is.null( dt$pagination$`next` ) ){
            break
        }
    
        current_page <- current_page + 1;
    }
    
    return(results)
    
}

# Create filters:
query_filters <- c(
    "areaType=nation",
    "areaName=England"
)

# Create the structure as a list or a list of lists:
query_structure <- list(
    date       = "date", 
    name       = "areaName", 
    code       = "areaCode", 
    daily     = "newCasesByPublishDate",
    cumulative = "cumCasesByPublishDate"
)

cases <- get_paginated_data(query_filters, query_structure)

list(
  "Shape"                = dim(cases),
  "Data (first 3 items)" = cases[0:3, 0:-1]
) -> report

print(report)

#Rename columns for better distinction
cases <- cases %>%
  rename(
    daily.cases.reported = daily,
    cumulative.cases.reported = cumulative
  )
```

## Merge the data frames of tests and cases

In the next step, we will merge both data frames by their **date column**. 

```{r}
tests.cases <- left_join(tests, cases, by = "date")
```

## Remove columns

The columns **name** and **code** are twice in the new data frame **tests.cases**. We will remove the duplicate columns as well as the cumulative cases as we don't need them for the positive testing rate. We also don't need the testing figures for the individual pillars as we have already created a column with the total tests which we will use for further calculations. 

```{r}
tests.cases <- tests.cases[-c(4:8,10,11,13)]
```

## Calculate the percentage of positive tests for each day 

```{r}
tests.cases$positive.testing.rate <- (tests.cases$daily.cases.reported/tests.cases$total.tests)*100
```

## Weekly positive testing rate

In a next step, we will calculate the weekly average for the positive testing rate. The **isoweek()** function from the lubridate package extracts the calendar week for each date in the data frame. 

The **isoweek()** package considers the ISO 8601 convention. If the week (starting on Monday) containing 1 January has four or more days in the new year, then it is considered week 1. Otherwise, it is the last week of the previous year, and the next week is week 1. 

```{r}
weekly.rates <- tests.cases %>% 
  group_by(week = isoweek(date)) %>%
  mutate(testing.rate.average = mean(positive.testing.rate))
```

## Keep every 7th row

As of now, there are still all days of the week in the data frame. We only need every 7th row, so that there is the weekly average (for the last seven days) displayed on the last day of the week. 

```{r}
weekly.rates.short <- weekly.rates %>% 
 filter(row_number() %% 7 == 1)
```

## Export

```{r}
write.csv(weekly.rates.short, file = "weekly_pos_testing_rate.csv")
write.csv(tests.cases, file = "positive_testing_rate.csv")
```
